---
title: "Mini Data-Analysis Deliverable 3"
author: "Armaghan Sarvar"
output: github_document
---


# Welcome to the last milestone in the mini data analysis project!

In Milestone 1, I explored my chosen data and came up with research questions. In Milestone 2, I obtained some results by making summary tables and graphs. 

In this (3rd) milestone, I will be sharpening some of the results obtained from the previous milestone by:

+ Manipulating special data types in R: factors and/or dates and times. 
+ Fitting a model object to the data, and extract a result.
+ Reading and writing data as separate files.

**NOTE**: It's possible that we might find the instructions too rigid for the data set. If this is the case, we may deviate from the instructions -- just should make sure we're demonstrating a wide range of tools and techniques taught in this class.

# Setup

Beginning by loading the `datateachr` and the `tidyverse` package below: 

```{r, message = FALSE}
library(datateachr) # <- might contain the data you picked!
library(tidyverse)
library(broom)
```

From Milestone 2, I chose two final research questions to focus on, which are the following: 

1. *Can we predict a diagnosis as malignant or benign based on the provided features from nuclei images?*
2. *How is the distribution of each feature value considering the samples being malignant or benign?*


# Exercise 1: Special Data Types

For this exercise, we will be choosing two of the three tasks below.

But first, tasks 1 and 2 below ask you to modify a plot you made in a previous milestone. The plot you choose should involve plotting across at least three groups (whether by facetting, or using an aesthetic like colour). Place this plot below (you're allowed to modify the plot if you'd like). If you don't have such a plot, you'll need to make one. Place the code for the plot below.

```{r}
cancer_sample
```

First, we extract the mean values of the feature `area_mean` for each of the `diagnosis` categories and put them in a new feature called `Mean_area_values`.

```{r}
mean_values <- cancer_sample %>%
  group_by(diagnosis)%>%
  summarise(Mean_area_values = mean(area_mean)) %>%
  mutate(across(where(is.numeric)))
mean_values
```
Now that we have a range of the the mean values, here, we group the numerical values of the `area_mean` column in the `cancer_sample` dataset regarding the above calculation.

```{r}
categorized <- cancer_sample %>%
     select(diagnosis, area_mean) %>%
     mutate(area = case_when (area_mean < 460 ~  "low", 
                              area_mean < 970 ~  "medium", 
                              TRUE ~ "high"))
categorized
```

Next, we can show a plot to visualize the number of samples in each of the obtained categories. 
```{r}
final_plot <- ggplot(categorized, aes(diagnosis, ..count..)) +
   geom_bar(fill= "blue", alpha = 0.3) +
  facet_wrap(~area)
final_plot
```
\

Now, let's choose two of the following tasks.

1. Produce a new plot that reorders a factor in the original plot, using the `forcats` package (3 points). Then, in a sentence or two, briefly explain why you chose this ordering (1 point here for demonstrating understanding of the reordering, and 1 point for demonstrating some justification for the reordering, which could be subtle or speculative.)

2. Produce a new plot that groups some factor levels together into an "other" category (or something similar), using the `forcats` package (3 points). Then, in a sentence or two, briefly explain why you chose this grouping (1 point here for demonstrating understanding of the grouping, and 1 point for demonstrating some justification for the grouping, which could be subtle or speculative.)

3. If the data has some sort of time-based column like a date (but something more granular than just a year):
    1. Make a new column that uses a function from the `lubridate` or `tsibble` package to modify the original time-based column. (3 points)
        - Note that you might first have to _make_ a time-based column using a function like `ymd()`, but this doesn't count.
        - Examples of something you might do here: extract the day of the year from a date, or extract the weekday, or let 24 hours elapse on the dates.
    2. Then, in a sentence or two, explain how the new column might be useful in exploring a research question. (1 point for demonstrating understanding of the function you used, and 1 point for the justification, which could be subtle or speculative).
        - For example, you could say something like "Investigating the day of the week might be insightful because penguins don't work on weekends, and so may respond differently".
    


**Task Number**: 1

```{r}
Task1_plot <- categorized %>%
  mutate(area=fct_relevel(area, "low", "medium", "high")) %>%
  ggplot(aes(diagnosis, ..count..)) +
  geom_bar(fill= "blue", alpha = 0.3) +
  facet_wrap(~area)
Task1_plot
```

\
As seen in the previous (initial) plot, the factors appeared alphabetically. Here, we change this so that they appear in ascending area_mean order using the `fct_relevel` function. This allows for an easier and more intuitive comparison based on this feature as now the plot cases are the mean of `area_mean` values in ascending order, for each diagnosis category. 


**Task Number**: 2

```{r}
Task1_plot <- categorized %>%
  mutate(area=fct_relevel(area, "low", "medium", "high")) %>%
  mutate(area = fct_collapse(area, other = c("medium", "high"))) %>%
  ggplot(aes(diagnosis, ..count..)) +
  geom_bar(fill= "blue", alpha = 0.3) +
  facet_wrap(~area)
Task1_plot
```
\

In the above plot, the two "medium" and "high" categories of the area column extracted before are considered as an independent group, and the "low" category has been demonstrated separately. This is helpful if we want to compare the "low" values with all other possible numbers of the `area-mean` features, and see how many benign and malignant samples are present for "low" area_mean vs. other area_mean values. In future analysis where the goal is to predict the diagnosis based on tumor features, this kind of plot is highly helpful when one needs to concentrate on specific data sample values and look at them separately to find the relationship of those samples with the final category. 
*Most importantly, this separation aggregates almost all the Malignant data into a single group ("other").*
    

# Exercise 2: Modelling

## 2.0

Here, we pick a research question, and pick a variable of interest (we'll call it "Y") that's relevant to the research question. Indicate these.


**Research Question**: Can we predict a diagnosis as malignant or benign based on the provided features from nuclei images?

**Variable of interest**: diagnosis

## 2.1

Fit a model or run a hypothesis test that provides insight on this variable with respect to the research question. Store the model object as a variable, and print its output to screen. We'll omit having to justify the choice, because we don't expect you to know about model specifics in STAT 545. 

- **Note**: It's OK if you don't know how these models/tests work. Here are some examples of things you can do here, but the sky's the limit. 
    - You could fit a model that makes predictions on Y using another variable, by using the `lm()` function.
    - You could test whether the mean of Y equals 0 using `t.test()`, or maybe the mean across two groups are different using `t.test()`, or maybe the mean across multiple groups are different using `anova()` (you may have to pivot the data for the latter two).
    - You could use `lm()` to test for significance of regression.



It should be noted that the `lm` function needs numeric `Y` values to fit a model with. However, the values of the `diagnosis` column are categorical. Hence, we use the following code which converts these categorical values to 0 and 1, so that the model is fit successfully. 

```{r}
categorical_diagnosis <- as.numeric(factor(as.matrix(cancer_sample[, "diagnosis"]))) - 1
new_df <- cancer_sample
new_df[, "diagnosis"] <- categorical_diagnosis
new_df
```

Also, the `ID` column does not have anything to do with the characteristics of the tumor! So, we drop this column.

```{r}
cancer_sample_final <- (select(new_df, -c(ID)))
cancer_sample_final
```


The following code fits the final linear model on the dataset that we created. The features consist of all columns except the `diagnosis` and `ID` columns. 

```{r}
fitted_model <- lm(diagnosis ~ ., data = cancer_sample_final)
fitted_model
```

We will also test the result for normalized dataset, as it was generated in the previous milestone.
```{r}
cancer_sample_final2 <- cancer_sample_final %>% mutate_at(2:30, funs((.-min(.))/max(.-min(.))))
cancer_sample_final2
```

```{r}
fitted_model2 <- lm(diagnosis ~ ., data = cancer_sample_final2)
fitted_model2
```

## 2.2

Produce something relevant from my fitted model: either predictions on Y, or a single value like a regression coefficient or a p-value. 

- Be sure to indicate in writing what you chose to produce. 
- Our code should either output a tibble (in which case you should indicate the column that contains the thing you're looking for), or the thing you're looking for itself.
- Obtain our results using the `broom` package if possible. If the model is not compatible with the broom function you're needing, then you can obtain the results by some other means, but first indicate which broom function is not compatible. 
  
Using the `glance` function from the broom package for better apprehension of the model:
```{r}
glance(fitted_model)
```


```{r}
glance(fitted_model2)
```

With the above code, we can see a tibble that contains the summary of p-values for hypothesis tests, the goodness of fit measures, and other useful information regarding our model which help us validate it. As seen, the `glance` function returns a one-row tibble.


Using the `augment` function from the broom package to add columns to the original data that was modeled.
```{r}
augment(fitted_model)
```


```{r}
augment(fitted_model2)
```

With the above code, we can add columns such as the predictions of the dependent variable using the model (`.fitted` column), residuals (`.resid` column), and cluster assignments to the original data.

Using the `tidy` function from the broom package to construct a tibble that summarizes the model’s statistical findings. 
```{r}
tidy(fitted_model)
```


```{r}
tidy(fitted_model2)
```


With the above code, we can have coefficients and p-values for each term in a regression, per-cluster information in clustering applications, or per-test information for multtest functions. As seen, the output of the `tidy` and `augment` functions is also a tibble.

# Exercise 3: Reading and writing data

Get set up for this exercise by making a folder called `output` in the top level of the project folder / repository. You'll be saving things there. 

## 3.1 (5 points)

Take a summary table that you made from Milestone 2 (Exercise 1.2), and write it as a csv file in the `output` folder. Use the `here::here()` function.

- **Robustness criteria**: We should be able to move the Mini Project repository / project folder to some other location on the computer, or move this very Rmd file to another location within the project repository / folder, and the code should still work.
- **Reproducibility criteria**: You should be able to delete the csv file, and remake it simply by knitting this Rmd file. 

Taking a summary table:
**Computing the range, mean, standard deviation and median of the "concavity_mean" variable across the groups of the "diagnosis" variable from the data.**

```{r}
summ_table <- (cancer_sample %>%
  group_by(diagnosis) %>%
  summarize(range(concavity_mean)[1], range(concavity_mean)[2], mean(concavity_mean), median(concavity_mean), sd(concavity_mean)))
summ_table
```
```{r}
dir.create(here::here("output"))
```

## Save the table to .csv file
```{r}
write_csv(summ_table, here::here("output", "summ_table.csv"))
```

## 3.2 (5 points)

Write the model object from Exercise 2 to an R binary file (an RDS), and load it again. Be sure to save the binary file in the `output` folder. Use the functions `saveRDS()` and `readRDS()`. 

- The same robustness and reproducibility criteria as in 3.1 apply here.

## Save the fitted model
```{r}
saveRDS(object = fitted_model, here::here("output", "fitted_model.rds"))
```

## Load the fitted model
```{r}
model_reloaded = readRDS(here::here("output", "fitted_model.rds"))
```

## Check if it was reloaded correctly
```{r}
model_reloaded %>% 
  tidy() %>% 
  knitr::kable(format = "markdown") 
```


# Tidy Repository

Now that this is the last milestone, our entire project repository should be organized. Here are the criteria we're looking for.

## Main README (3 points)

There should be a file named `README.md` at the top level of the repository. Its contents should automatically appear when you visit the repository on GitHub. 

Minimum contents of the README file:

- In a sentence or two, explains what this repository is, so that future-you or someone else stumbling on the repository can be oriented to the repository.
- In a sentence or two (or more??), briefly explains how to engage with the repository. You can assume the person reading knows the material from STAT 545A. Basically, if a visitor to the repository wants to explore the project, what should they know?

Once you get in the habit of making README files, and seeing more README files in other projects, you'll wonder how you ever got by without them! They are tremendously helpful.

## File and Folder structure (3 points)

You should have at least four folders in the top level of the repository: one for each milestone, and one output folder. If there are any other folders, these are explained in the main README.

Each milestone document is contained in its respective folder, and nowhere else. 

Every level-1 folder (that is, the ones stored in the top level, like "Milestone1" and "output") has a `README` file, explaining in a sentence or two what is in the folder, in plain language (it's enough to say something like "This folder contains the source for Milestone 1").

## Output (2 points)

All output is recent and relevant: 

- All Rmd files have been `knit`ted to their output, and all data files saved from Exercise 3 above appear in the `output` folder. 
- All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
- There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted. 

Our recommendation: delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant. 

PS: there's a way where you can run all project code using a single command, instead of clicking "knit" three times. More on this in STAT 545B!

## Error-free code (1 point)

This Milestone 3 document knits error-free. (We've already graded this aspect for Milestone 1 and 2)

## Tagged release (1 point)

You've tagged a release for Milestone 3. (We've already graded this aspect for Milestone 1 and 2)