---
title: "Mini Data-Analysis Milestone 1"
author: "Armaghan Sarvar"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the the Necessary Packages

```{r}
library(datateachr)  # provides 7 semi-tidy datasets
library(tidyverse)  # provides data analysis libraries
```


## 1.1 Choosing four datasets from the `datateachr` package

Here, the four datasets that I have chosen for analysi are listed and briefly described.

1. `cancer_sample`: Acquired courtesy of UCI Machine Learning Repository. It is reported to have 569 rows and 32 columns.
2. `steam_games`: Acquired courtesy of Kaggle. It is reported to have 40833 rows and 21 columns.
3. `vancouver_trees`: Acquired courtesy of The City of Vancouver’s Open Data Portal. It is reported to have 146611 rows and 20 columns.
4. `apt_buildings`: Acquired courtesy of The City of Toronto’s Open Data Portal. It is reported to have 3455 rows and 37 columns.


## 1.2 In the following parts, the four chosen datasets will be explored. 
More specifically, the `dplyr` package will be used to help us find out at least 3 attributes about each dataset. 

### Dataset 1

```{r}
cancer_sample
```

* Class Type
```{r}
class(cancer_sample)
```


* Dimension of the dataset
```{r}
dim(cancer_sample)
```


* Variables (features) of the dataset
```{r}
names(cancer_sample)
```

* Overall view of the data frame

```{r}
summary(cancer_sample)
```


* See the unique values in the "diagnosis" column
```{r}
unique(cancer_sample$diagnosis)
```


* Counting the number of samples for each diagnosis result
```{r}
cancer_sample %>%
  group_by(diagnosis) %>%
  summarize(n())
```


### Dataset 2


```{r}
steam_games
```

* Class Type
```{r}
class(steam_games)
```


* Dimension of the dataset
```{r}
dim(steam_games)
```


* Variables (features) of the dataset
```{r}
names(steam_games)
```


* Overall view of the data frame

```{r}
summary(steam_games)
```


* See the unique values in the "types" column
```{r}
unique(steam_games$types)
```


* Counting the number of samples for each type reported
```{r}
steam_games %>%
  group_by(types) %>%
  summarize(n())
```



### Dataset 3


```{r}
vancouver_trees
```

* Class Type
```{r}
class(vancouver_trees)
```


* Dimension of the dataset
```{r}
dim(vancouver_trees)
```


* Variables (features) of the dataset
```{r}
names(vancouver_trees)
```


* Overall view of the data frame

```{r}
summary(vancouver_trees)
```


* See the unique values in the "street_side_name" column
```{r}
unique(vancouver_trees$street_side_name)
```


* Counting the number of samples for each street_side_name reported
```{r}
vancouver_trees %>%
  group_by(street_side_name) %>%
  summarize(n())
```


### Dataset 4


```{r}
apt_buildings
```

* Class Type
```{r}
class(apt_buildings)
```


* Dimension of the dataset
```{r}
dim(apt_buildings)
```


* Variables (features) of the dataset
```{r}
names(apt_buildings)
```


* Overall view of the data frame

```{r}
summary(apt_buildings)
```


* See the unique values in the "heating_type" column
```{r}
unique(apt_buildings$heating_type)
```


* Counting the number of samples for each heating_type reported
```{r}
apt_buildings %>%
  group_by(heating_type) %>%
  summarize(n())
```


## 1.3 Narrowing down top 2 datasets
Based on the exploration done in the previous part (section 1.2), I found `cancer_sample` and `steam_games` most fascinating ones.

* `cancer_samples`: 
1. First of all, I find the application and usage of this dataset in healthcare highly interesting and important. 
By analyzing this dataset, we can predict whether a possible detected tumour is malignant or benign based on its features.
2. As the code is shown in the previous section, the number of samples for each diagnosis result (357	and	212) are close to each other, which says we have a balanced dataset.
3. We do not have any NaN values in the diagnosis column.
4. Many quantitative values are present in the dataset which makes statistical analysis in terms of relation between different values more meaningful.

* `steam_games`: 
1. In this dataset, features and their values are human-readable and no expertise is needed for understanding them.
2. The application of marketing analysis that can be performed on it is highly interesting.
3. As the code is shown in the previous section, the number of samples for each of the "app" and "bundle" type values is high enough which can help us analyze the differences between the two, for example based on the reviews recorded.  
4. Dataset has a "recent_reviews" column, which enables many research questions about the data. In the following, the possible analysis is shown:

```{r}
nrow(steam_games %>%
  filter(str_detect(recent_reviews, "Positive")))
```
```{r}
nrow(steam_games %>%
  filter(str_detect(recent_reviews, "Negative")))
```

```{r}
nrow(steam_games %>%
  filter(str_detect(recent_reviews, "Mixed")))
```



## 1.4 Choosing the final dataset

My final choice in dataset selection would be `cancer_sample`.
A possible research question I'm interested in answering is: 
** How can we diagnose cancer having different quantitative variables and characteristics of the tomour masses? **



# Task 2: Exploring the chosen dataset

## 2.1 
In this section, 4 exercises will be conducted to help me dive deeper into the data. These exercises are listed below:
1. Plotting the hostogram of the numeric variable "perimeter_mean":


```{r}
cancer_sample %>%
  ggplot(aes(x=perimeter_mean)) + 
  geom_histogram(aes(y=..density..), bins=20, colour="blue", fill="cyan") +
  geom_density(fill="grey", alpha=0.5)
```
As we can in the above plot, the frequencies of the "perimeter_mean" column values form a left-skewed curve. 

```{r}
ggplot(cancer_sample) +
    geom_jitter(aes(x = perimeter_mean, y = diagnosis), colour="blue", alpha = 0.1)
```
As shown in the above plot, we can have a separation of Benign and Malignant diagnosis values based on the values of the "perimeter_mean" feature.


